### 2023-10-18 Demo_marrow_facs.Rmd
benchmark&figure_code

---
title: "Tutorial for Runing Mouse Bone Marrow Dataset"
author: "Ruiqi Li"
date: '2023-10-12'
output: pdf_document
bibliography: references.bib
---

# Install & Load Packages
```{r}
# Install necessary tools if not already present
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
# Define package lists
bioc_packages <- c("Seurat")
github_packages <- c("satijalab/SeuratWrappers")

# Install packages
sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))

library("Seurat")
library("SeuratWrappers")
```

# Preparing input data
```{r}
# If load url doesn't work, please download it and load from your working directory
load(url("https://figshare.com/ndownloader/files/13092380"))
tiss <- UpdateSeuratObject(tiss)
```
We obtain the following objects from this Seurat Obj:
* `feature_space`: The output of PCA. This matrix contains the first 20 pricipal components (PCs).
* `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
* `dat`: a matrix object with the log-normalized expression of genes (rows) in each cell (columns)
* `cell_label`: Any metadata you wish to visualize and examine related to cells (here we take the cell type annotation)
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class
```
We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)
```{r}
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

# Running LMD Step by Step
This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'
```{r}
source("./LMD_function.R", echo = F)
```
## Step1: Build KNN Graph
```{r}
# Construct knn graph
res = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)

A = res$adj_matrix # Adjacency Matrix
W = res$graph # Symmetrized Graph ((A + AT) / 2)

# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```

## Step2: Diffuse Genes on the cell graph
```{r}
# Generate $P^{t}, t = 2,4,8,...,max_time$
P_ls = Obtain_Pls(W = W, max_time = 2^10)
# Generate initial gene distribution
rho = Rowwise_normalize(dat[,colnames(W)])
# Check One Gene Diffusion on Cell graph by time
gene = "Fcnb"
print(FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2,16,128),gene_name = gene))
```

## Step3: Obtain Diffusion KL score & Cumulative score
```{r}
# Obtain diffusion KL score
res = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)

# Check Diffusion KL pattern for several genes
genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes_label = rep(c("good","bad"),each = 3)
Visualize_score_pattern(res, genes = genes, label_class = genes_label, facet_class = NULL)

# Show result
show_result_lmd(res,n = 10)
```

# Running LMD in one step
```{r}
res = LMD(dat,feature_space)
```

# Identify Gene Modules
## Step1: Deriving Gene Modules
```{r}
# Select top1000 genes as local genes
local_gene = show_result_lmd(res, n = 1000)
local_gene = rownames(local_gene)

# Denoising
#' We use ALRA @linderman2022zero to Denoise our data
tiss <- RunALRA(tiss, assay = "RNA")
DefaultAssay(tiss) <- "RNA"
dat_alra = as.matrix(tiss[["alra"]]@data)[local_gene,]

# Calculate Gene-gene co-localization distance
distance_method = "jaccard"
dist = Calculate_distance(dat_alra, method = distance_method)

# Hierarchical Tree
clustering_method = "average"
gene_hree = hclust(dist, method = clustering_method)

# Obtain Gene modules
gene_partition = dynamicTreeCut::cutreeDynamic(
  dendro = gene_hree, 
  distM = as.matrix(dist), deepSplit = 2,
  pamStage = TRUE, pamRespectsDendro = TRUE,
  minClusterSize = 5)
names(gene_partition) = labels(dist)

# Re-name gene modules based on the order of htree
gene_partition = as.factor(gene_partition)
module_order = order(sapply(levels(gene_partition), function(mod) {
  median(which(gene_partition[gene_hree$order] == mod))
}))
levels(gene_partition)[module_order] = seq(nlevels(gene_partition))
gene_partition = factor(gene_partition,levels = seq(nlevels(gene_partition)))
rm(module_order)
```
### Visualize Heatmap
```{r}
p = pheatmap(as.matrix(dist),
              annotation_col = data.frame(Module = gene_partition), 
              cluster_rows = gene_hree, 
              cluster_cols = gene_hree,
              ann_colors = list(Module = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))),
              treeheight_row = 0,
              col=colorRampPalette(c("red","white","blue4"))(99),
              show_colnames = FALSE, 
              show_rownames = FALSE,
              main = paste(distance_method,
                           clustering_method,sep = "-"),
              silent = TRUE )
as.ggplot(p)
```

## Step2: Computing Per-Cell Module Activity Scores
```{r}
knn = 10; major_vote = 5
A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
cell_block_loop = lapply(1:100,function(loop){
  cat("\n",loop,"\n")
  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
})
cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
cell_block = cell_block_prop
```
### Visualize Module on Cell embeddings
```{r}
pl = lapply(levels(gene_partition),function(module){
  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
})
plot_grid(plotlist = pl,ncol = 10) # may takes long time to plot
```

## Step3: Identify Celltype-specific activated Modules
```{r}
# Create one-hot vector of Cell Type
one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))

# Order Cell Type based on its centroid in Featurespace
Idents(tiss) = cell_label
ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]

# Calculate Jaccard index between module activity & cell type
jaccard_index <- apply(cell_block, 2, function(module_col) {
  apply(one_hot_matrix, 2, function(one_hot_col) {
    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard")
  })
})
jaccard_index = jaccard_index[rev(ordered_cell_types),]
rm(ordered_cell_types)
```

### Visualize Heatmap of Jaccard Index Matrix
```{r}
pheatmap(jaccard_index, 
         cluster_rows=FALSE, cluster_cols=FALSE, 
         scale = "none", 
         color = colorRampPalette(rev(c("firebrick3", "white", "navy")))(100))
```

### Visualize How CellType-Module is Matched on Cell Embeddings
For each cell type, select the most specific modules for visualization (maximum jaccard index)
```{r}
tiss@meta.data = cbind(tiss@meta.data,one_hot_matrix)
jaccard_thred = 0.4

# Get Module plot
pl_module = lapply(levels(gene_partition),function(module){
  FeaturePlot_custom(value = cell_block[,module], coord = visual_space, value_name = "Module Score",title_name = sprintf("Module %s (%d genes)",module, sum(gene_partition == module)))
})
names(pl_module) = levels(gene_partition)

# Get CellType One-hot plot
pl_ct = lapply(colnames(one_hot_matrix),function(ct){
  FeaturePlot_custom(value = as.factor(one_hot_matrix[,ct]), coord = visual_space, value_name = "Indicator",title_name = ct)
})
names(pl_ct) = colnames(one_hot_matrix)

pl <- lapply(rownames(jaccard_index)[rev(order(apply(jaccard_index,1,max)))], function(celltype){
  if(max(jaccard_index[celltype,]) > jaccard_thred){
    p1 = pl_ct[[celltype]]
    module_name = colnames(jaccard_index)[which.max(jaccard_index[celltype,])]
    p2 = pl_module[[module_name]]
    p = as.ggplot(p1) + as.ggplot(p2) + plot_annotation(paste0("Jaccard Index: ",round(jaccard_index[celltype,module_name],2)))
    p
  }
})
```
### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
```{r}
module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x > 0.4)),collapse = ",")})
```

## Save Intermediate data
```{r}
save(gene_partition,dist,gene_hree,cell_block,jaccard_index,file = "./marrow_result/module_result.rdata")
```

## Step4: Module Enrichment Analysis
### GO enrichment Analysis
```{r}
if (!requireNamespace("clusterProfiler", quietly = TRUE)) BiocManager::install("clusterProfiler")
library(clusterProfiler)

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))

# Filter GO terms based on p values < 0.05
ego_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichGO(gene = names(gene_partition)[gene_partition == i],
           OrgDb = 'org.Mm.eg.db', # mouse
           keyType = "SYMBOL",
           ont = "ALL",
           pAdjustMethod = "BH",
           universe = universe_df$symbol,
           pvalueCutoff = 0.05)
})

# Only keep modules which has at lease one GO term having adjusted p-value < 0.05
top_go_clusters <- which(sapply(ego_result, function(x)
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 GO terms
top_enrich_go <- lapply(top_go_clusters, function(i) ego_result[[i]]@result[1:5,])
```

### Reactome enrichment Analysis
```{r}
if (!requireNamespace("ReactomePA", quietly = TRUE)) BiocManager::install("ReactomePA")
if (!requireNamespace("org.Mm.eg.db", quietly = TRUE)) BiocManager::install("org.Mm.eg.db")
library("clusterProfiler")
library("ReactomePA")
library("org.Mm.eg.db")

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
universe_df = universe_df[!is.na(universe_df$entrez),]

# Filter Pathway based on p values < 0.05
epathway_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
                organism = "mouse",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05,
                universe = universe_df$entrez)
})

# Only keep modules which has at lease one Pathway having adjusted p-value < 0.05
top_pathway_clusters <- which(sapply(epathway_result, function(x) 
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 Pathway terms
top_enrich_pathway <- lapply(top_pathway_clusters, function(i) epathway_result[[i]]@result[1:5,])
```

## Step5: Generate Gene Module Annotation Table
```{r}
df_module = data.frame(Module = levels(gene_partition))
df_module[,"Gene"] = unlist(lapply(levels(gene_partition),function(x){paste(names(gene_partition)[gene_partition == x],collapse = ",")}))
df_module[,"celltype"] = module_celltype[levels(gene_partition)]
df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x) paste(x[['Description']],collapse = ",")))
df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x) paste(x[['Description']],collapse = ",")))
df_module[is.na(df_module)] = ""
write.csv(df_module,file = "module_description.csv",row.names = FALSE)
```