### 2023-10-27 Marker_identfication_benchmark.Rmd

---
title: "Marker_identification_Benchmark"
output: pdf_document
date: '2023-10-16'
---
# Load Packages
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("BiocManager", quietly = TRUE)) BiocManager::install("Seurat")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")

library("Seurat")
library("dplyr")
```

# Load Data & Process
## Load Data
```{r}
path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
tissue_name = "marrow_facs" #lung_facs, pancreas_facs
tiss <- readRDS(paste0(path,tissue_name,".rds"))
```
## Process
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class

Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
selected_genes = names(selected_genes)[selected_genes]
dat = dat[selected_genes,,drop = FALSE]
raw_count = as.matrix(tiss@assays$RNA@counts)[selected_genes,,drop = FALSE]
total_count = tiss@meta.data$nCount_RNA
```

# Define Ground truth Marker
## Criterion1: Fold-change
For each gene, sorting cell types by mean expression for each gene and computing fold changes between consecutive types, take the maximum value among the N-1 fold change values, given N cell types.
```{r}
path = "./benchmark_result/"
tissue_name = "marrow_facs"
```

```{r}
avg_exp = AverageExpression(subset(tiss,features = selected_genes), assays = "RNA", slot = "counts", group.by = "cell_ontology_class") %>% as.data.frame()

avg_exp_ordered <- avg_exp %>%
  rowwise() %>%
  mutate(cell_order_val = list(sort(c_across(everything()), decreasing = FALSE))) %>%
  ungroup()

avg_exp_ordered = avg_exp_ordered %>%
  rowwise() %>%
  mutate(log_fold_change = list(c(NA, diff(log2(unlist(cell_order_val)+1))))) %>%
  ungroup()

max_logfc = unlist( lapply(avg_exp_ordered$log_fold_change,function(vec){max(vec,na.rm = TRUE)}) )
names(max_logfc) = rownames(avg_exp)

max_logfc = sort(max_logfc,decreasing = TRUE)
write.table(max_logfc,file = paste0(path,tissue_name,"_ground_truth_c1.txt"))
```
## Criterion2: CellMarkerDB
Downloaded marker from http://bio-bigdata.hrbmu.edu.cn/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
<!-- PanglaoDB: https://panglaodb.se/markers/PanglaoDB_markers_27_Mar_2020.tsv.gz and exclude genes that reported only once -->
```{r}
# Load CellMarkerDB
cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) %>% group_by(across(everything())) %>% dplyr::summarise(count = n(), .groups = "drop") %>% filter(count >= 1) # Lung, Pancreas

# Match gene name
celldb_marker = union(cell_marker_db$gene_name1,cell_marker_db$gene_name2)
celldb_marker = unlist( lapply(celldb_marker,function(i){
  id = grep(paste0("^", i, "$"), selected_genes, ignore.case = TRUE)
  if(length(id)==0){return(NA)}
  else{return(selected_genes[id])}
}) )
celldb_marker = celldb_marker[!is.na(celldb_marker)] %>% unique()
write.table(celldb_marker,file = paste0(path,tissue_name,"_ground_truth_c2.txt"))
```

# Run Each Method and save results
```{r}
path = "./benchmark_result/"
# tissue_name = "marrow_facs"
```

## LocalMarkerDetector
Genes were ranked by the increasing order of Cumulative Diffusion-KL score
```{r}
method = "lmd"
source("./LMD_function.R", echo = F)
res = LMD(dat,feature_space,max_time = 2^10,knn = 7)
marker = show_result_lmd(res, n = length(selected_genes))
marker[order(marker$score),'rank'] = 1:nrow(marker)
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Seurat v3 Wilcoxon unfiltered
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)
method = "wilcox_no_filter"
marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## Seurat v3 default
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)

method = "wilcox_default"
marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0.25,min.pct = 0.1)

gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## singleCellHaystack
https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
Genes were ranked based on the increasing order of log(adjusted P-value)
```{r}
method = "haystack"
library(singleCellHaystack)
set.seed(123)
res <- haystack(x = feature_space, expression = dat)
marker <- show_result_haystack(res.haystack = res)
marker[order(marker$log.p.adj),"rank"] = 1:nrow(marker)

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```
## SEMITONES
https://github.com/ohlerlab/SEMITONES/tree/master
https://github.com/ohlerlab/SEMITONES_paper/blob/main/notebooks/10_SEMITONES_benchmark_corrected.ipynb

Genes were ranked by their lowest adjusted P-value in any reference cell (increasing order), if two genes had the same p.adj, prioritize genes with larger absolute enrichment score in any reference cell.
```{python}
import SEMITONES
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
from SEMITONES.cell_selection import from_knn_dist
from SEMITONES.cell_selection import from_gui
from SEMITONES.cell_selection import from_2D_embedding
from SEMITONES.cell_selection import get_cells_from_gui
from sklearn.metrics.pairwise import pairwise_kernels
from SEMITONES.enrichment_scoring import calculate_escores
from SEMITONES.enrichment_scoring import permute
from SEMITONES.enrichment_scoring import sig_interval
from SEMITONES.support_funcs import pairwise_similarities
from SEMITONES.support_funcs import sig_dictionary
from SEMITONES.support_funcs import sig_bool
from SEMITONES.enrichment_scoring import pvals_per_cell

# Load data
data = pd.DataFrame(r.dat.T)  # counts
PC20 = pd.DataFrame(r.feature_space)
# Reference Cell
g = 8.6e-4  
S = pairwise_kernels(PC20, metric="rbf", gamma=g) 
median = np.median(S, axis=0)
start = int(np.argmin(median))
num_refcell = int(data.shape[1]*0.005)
dd_rcells = from_knn_dist(X=PC20, 
                          n_ret=num_refcell,
                          start=start, 
                          metric="rbf",
                          metric_params={"gamma": g})
S = pairwise_similarities(PC20,
                          query=dd_rcells,
                          metric="rbf",
                          metric_params={"gamma": g})
dd_rcells = data.index[dd_rcells]
escores = calculate_escores(data.values, query=dd_rcells, S=S)

# semi_ranks = escores.abs().rank(ascending=False, method="dense").min(1)

P = permute(data.values)
pscores = calculate_escores(P, query=dd_rcells, S=S)
pvals = pvals_per_cell(escores,pscores,ret = "q")
# semi_ranks = pvals.min(1).rank(ascending=True, method="dense")

result = pd.DataFrame({'escore': escores.abs().max(1), 'padj': pvals.min(1)})
```
```{r}
method = "semi"
marker = reticulate::py$result
rownames(marker) = rownames(dat)
marker = marker %>% arrange(.,padj,desc(escore))
marker$'rank' = 1:nrow(marker)

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Hotspot
https://hotspot.readthedocs.io/en/latest/CD4_Tutorial.html
Genes were ranked by the increasing order FDR value (if two genes had the same, prioritize genes with larger Z-score).
```{r}
cell_id = colnames(raw_count)
gene_id = rownames(raw_count)
```
```{python}
import hotspot
import scanpy as sc
import pandas as pd
from scipy.sparse import csc_matrix
adata = sc.AnnData(X = r.raw_count.T,obs = pd.DataFrame(index = r.cell_id), var = pd.DataFrame(index = r.gene_id))
adata.obs['total_counts'] = r.total_count
adata.layers["counts"] = adata.X.copy()
adata.obsm['X_pca'] = r.feature_space
adata.layers["counts_csc"] = csc_matrix(adata.layers["counts"])

hs = hotspot.Hotspot(
    adata,
    layer_key="counts_csc",
    model='danb',
    latent_obsm_key="X_pca",
    umi_counts_obs_key="total_counts"
)
hs.create_knn_graph(weighted_graph=False, n_neighbors=30)

# Determining informative genes
hs_results = hs.compute_autocorrelations()
```
```{r}
method = "hotspot"
marker = reticulate::py$hs_results
marker[order(marker$FDR),"rank"] = 1:nrow(marker)
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Marcopolo
The rank of genes is given by `Marcopolo_rank`
```{python}
import pickle
import numpy as np
import pandas as pd
import torch
import anndata as ad
import scanpy as sc
import matplotlib.pyplot as plt
import MarcoPolo

adata = sc.AnnData(X = r.raw_count.T)

if "size_factor" not in adata.obs.columns:
    norm_factor = sc.pp.normalize_total(adata, exclude_highly_expressed=True, max_fraction= 0.2, inplace=False)["norm_factor"]
    adata.obs["size_factor"] = norm_factor/norm_factor.mean()
    print("size factor was calculated")

regression_result = MarcoPolo.run_regression(adata=adata, size_factor_key="size_factor",
                         num_threads=1, device="cpu")

result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
```
```{r}
method = "marcopolo"
marker = reticulate::py$result
marker$'rank' = marker$MarcoPolo_rank + 1
write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = TRUE)
```

## Highly Variable Genes
```{r}
method = "hvg"

DefaultAssay(tiss) <- "RNA"
# tiss <- NormalizeData(tiss, normalization.method = "LogNormalize", scale.factor = 1e6)
marker <- subset(tiss,features = selected_genes) %>% ScaleData(features = selected_genes) %>% 
  FindVariableFeatures(nfeatures = 5000) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())

write.table(marker,file = sprintf("%s%s_%s.csv",path,method,tissue_name),row.names = FALSE)
```

# Integrate rank from all methods into one dataframe
```{r}
path = "./benchmark_result/"
# tissue_name = "marrow_facs"
```
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)

df_benchmark = do.call(cbind,lapply(method_ls, function(method){
  marker = read.table(sprintf("%s%s_%s.csv",path,method,tissue_name),header = TRUE)
  if("gene" %in% colnames(marker)){
    return(marker$rank[match(selected_genes,marker$gene)])
  }else{
    return(marker$rank[match(selected_genes,rownames(marker))])
  }
}))
df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)
df_benchmark = data.frame(df_benchmark,row.names = selected_genes)
colnames(df_benchmark) = method_ls

write.table(df_benchmark,file = paste0(path,tissue_name,"_benchmark_rank_table.csv"),row.names = TRUE)
```

# AUROC
```{r}
path = "./benchmark_result/"
# tissue_name = "marrow_facs"
df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))

method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
method_fullname = setNames(c("LMD","singleCellHaystack","Hotspot","Marcopolo","SEMITONES","Seurat v3 default", "Seurat v3 Wilcoxon unfiltered","HighlyVariableGenes"),method_ls)
```
```{r}
max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]
gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

# AUC table
df = do.call(rbind,lapply(1:length(gt_list),function(i){
  true_marker = gt_list[[i]]
  df_benchmark$"gt" = 0
  df_benchmark[true_marker,"gt"] = 1

  library(pROC)
  roc_df = do.call(rbind,lapply(setdiff(colnames(df_benchmark),"gt"), function(method){
    roc = roc(df_benchmark$gt, df_benchmark[,method], direction = ">")
    data.frame(FPR = 1 - roc$specificities, TPR = roc$sensitivities, Method = method, AUROC = as.numeric(auc(roc)))
  }) )

  df = roc_df %>% select(c("Method","AUROC")) %>% distinct()
  df$'gt_set' = names(gt_list)[i]
  df
}))
df$Method = as.factor(df$Method)
levels(df$Method) = method_fullname[levels(df$Method)]
df$gt_set = factor(df$gt_set,levels = names(gt_list))
write.table(df,file = paste0(path,tissue_name,"_auroc.csv"),row.names = FALSE)
```

```{r}
# Concordance table
do.call(cbind,lapply(gt_list[c(2,21)],function(true_marker){
  df_benchmark$"gt" = 0
  df_benchmark[true_marker,"gt"] = 1
  apply(df_benchmark,2,function(x){
    sum(df_benchmark[order(x),"gt"][1:100])
  })
}))

# AUC curve
p = ggplot(roc_df, aes(x = FPR, y = TPR, color = label)) +
  geom_line(size = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + labs(title = sprintf("CellMarkerDB (%d genes)",length(true_marker)),
     x = "False Positive Rate",
     y = "True Positive Rate") +  scale_colour_manual(values= coldef)
```

# Density Index
```{r}
path = "./benchmark_result/"
# tissue_name = "marrow_facs"
```
```{r}
df_benchmark = read.table(paste0(path,tissue_name,"_benchmark_rank_table.csv"))

score_measure =do.call(rbind,lapply(c(method_ls,"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    top_n = c(seq(100,900,100),seq(1000,5000,1000))
  }
  result = unlist(lapply(top_n, function(top){
    tiss <- RunPCA(tiss, features = top_gs[1:top], npcs = 20)
    d = dist(Embeddings(tiss, reduction = "pca"))
    c = as.integer(factor(tiss$cell_ontology_class[labels(d)]))

    # # Silhouette score
    # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
    #   group_by(cluster) %>%
    #   summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # Density-index
    nn.dist = FNN::get.knn(Embeddings(tiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(tiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2))
    s2 = length_scale/mean(nn.dist)

    s2

  }))
  # rownames(result) = c("Mean_SI","DI")
  result = data.frame(TopGene = top_n,Method = method,DensityIndex = result)
}) )

write.table(score_measure,file = paste0(path,tissue_name,"_density_index.csv"),row.names = FALSE)
```

# Sensitivity Test (Try different K)
```{r}
path = "./marrow_result/"
# tissue_name = "marrow_facs"
```
```{r}
# RunLMD under differen K
source("./LMD_function.R", echo = F)
score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
  try({
    res = LMD(dat,feature_space,knn = kNN)
    return(res$cumulative_score)
  },silent = TRUE)
}))
colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
write.table(score_df,file = paste0(path,tissue_name,"_different_knn.csv"),row.names = FALSE)
```
```{r}
# Obtain AUROC
score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))
path = "./benchmark_result/"
max_logfc = read.table(paste0(path,tissue_name,"_ground_truth_c1.txt")) %>% rownames()
celldb_marker = read.table(paste0(path,tissue_name,"_ground_truth_c2.txt"))[,1]

gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

auc_df = apply(score_ranked,2,function(x){
  unlist(lapply(gt_list,function(genes){
    roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
    pROC::auc(roc)
  }))
})
path = "./marrow_result/"
auc_df = data.frame(auc_df,gt = names(gt_list))
write.table(auc_df,file = paste0(path,tissue_name,"_sensitivity_auroc.csv"),row.names = FALSE)
```