### 2023-10-29 Demo_marrow_facs.Rmd
fig-adjustment

---
title: "Tutorial for Runing Mouse Bone Marrow Dataset"
author: "Ruiqi Li"
date: '2023-10-12'
output: pdf_document
bibliography: references.bib
---

# Install & Load Packages

```{r}
# Install necessary tools if not already present
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("remotes", quietly = TRUE)) install.packages("remotes")
# Define package lists
bioc_packages <- c("Seurat",
                   "clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2")
github_packages <- c("satijalab/seurat-wrappers")
# Note: repeating packages from LMD_function.R as knitr does not install them properly
cran_packages <- c("igraph","ggplot2", "cowplot", "RColorBrewer",
                   "data.table", "dplyr", "patchwork",
                   "pheatmap", "ggplotify", "ggraph",
                   "ClusterR", "Rcpp", "RcppArmadillo", "tictoc")

# Install packages
sapply(cran_packages, function(pkg) if(!requireNamespace(pkg, quietly = TRUE)){install.packages(pkg)})
sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
sapply(github_packages, function(pkg) if (!requireNamespace(strsplit(pkg, "/")[[1]][2], quietly = TRUE)) remotes::install_github(pkg))

library("Seurat")
library("SeuratWrappers")
library("pheatmap")
library("ggplot2")
library("ggplotify")
```

# Preparing input data

```{r}
if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
 }
load('facs_Marrow_seurat_tiss.Robj')
tiss <- UpdateSeuratObject(tiss)
```
We obtain the following objects from this Seurat Obj:
* `feature_space`: The output of PCA. This matrix contains the first 20 principal components (PCs).
* `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
* `dat`: a matrix object with the log-normalized expression of genes (rows) in each cell (columns)
* `cell_label`: Any metadata you wish to visualize and examine related to cells (here we take the cell type annotation)
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class
```

We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)

```{r}
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

# Running LMD Step by Step

This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'

```{r}
source("./LMD_function.R", echo = F)
```

## Step1: Build KNN Graph

```{r, fig.width=20}
# Construct knn graph
knn_result = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)

A = knn_result$adj_matrix # Adjacency Matrix
W = knn_result$graph # Symmetrized Graph ((A + AT) / 2)

# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```

## Step2: Diffuse Genes on the cell graph

```{r, fig.width=20}
# Generate $P^{t}, t = 2,4,8,...,max_time$
P_ls = Obtain_Pls(W = W, max_time = 2^10)
# Generate initial gene distribution
rho = Rowwise_normalize(dat[,colnames(W)])
# Check One Gene Diffusion on Cell graph by time
gene = "Fcnb"
print(FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2,16,128),gene_name = gene))
```

## Step3: Obtain Diffusion KL score & Cumulative score

```{r}
# Obtain diffusion KL score
score_result = fast_calculate_multi_score(W = W, init_state = rho, P_ls = P_ls)

# Check Diffusion KL pattern for several genes
genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes_label = rep(c("good","bad"),each = 3)
Visualize_score_pattern(score_result, genes = genes, label_class = genes_label, facet_class = NULL)

# Show result
show_result_lmd(score_result,n = 10)
```

# Running LMD in one step

```{r}
lmd_result = LMD(dat, feature_space)
```

# Identify Gene Modules

## Step1: Deriving Gene Modules
```{r}
# Select top1000 genes as local genes
local_gene = show_result_lmd(lmd_result, n = 1000)
local_gene = rownames(local_gene)

# Denoising
#' We use ALRA @linderman2022zero to Denoise our data
tiss <- RunALRA(tiss, assay = "RNA")
DefaultAssay(tiss) <- "RNA"
dat_alra = as.matrix(tiss[["alra"]]@data)[local_gene,]

# Calculate Gene-gene co-localization distance
distance_method = "jaccard"
dist = Calculate_distance(dat_alra, method = distance_method)

# Hierarchical Tree
clustering_method = "average"
gene_hree = hclust(dist, method = clustering_method)

# Obtain Gene modules
gene_partition = dynamicTreeCut::cutreeDynamic(
  dendro = gene_hree, 
  distM = as.matrix(dist), deepSplit = 2,
  pamStage = TRUE, pamRespectsDendro = TRUE,
  minClusterSize = 5)
names(gene_partition) = labels(dist)

# Re-name gene modules based on the order of htree
gene_partition = as.factor(gene_partition)
module_order = order(sapply(levels(gene_partition), function(mod) {
  median(which(gene_partition[gene_hree$order] == mod))
}))
levels(gene_partition)[module_order] = seq(nlevels(gene_partition))
gene_partition = factor(gene_partition,levels = seq(nlevels(gene_partition)))
rm(module_order)
```

### Visualize Heatmap
```{r}
p = pheatmap(as.matrix(dist),
              annotation_col = data.frame(Module = gene_partition), 
              cluster_rows = gene_hree, 
              cluster_cols = gene_hree,
              ann_colors = list(Module = setNames(colorRampPalette(brewer.pal(12, "Paired"))(nlevels(gene_partition)),levels(gene_partition))),
              treeheight_row = 0,
              col=colorRampPalette(c("firebrick3", "white"))(99),
              show_colnames = FALSE,
              show_rownames = FALSE,
              main = paste(distance_method,
                           clustering_method,sep = "-"),
              silent = TRUE )
as.ggplot(p)
```

## Step2: Computing Per-Cell Module Activity Scores
```{r}
knn = 10; major_vote = 5
A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space)$'adj_matrix'
cell_block_loop = lapply(1:100,function(loop){
  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat, cell_kNN_graph = A, major_vote)
})
cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
cell_block = cell_block_prop
```

## Step3: Identify Celltype-specific activated Modules
```{r}
cell_label = as.factor(tiss$cell_ontology_class)
abbrev = levels(cell_label)
abbrev[grepl("T cell",abbrev)] = "T cell"
abbrev[grepl("natural killer",abbrev)] = "NK cell"
abbrev[abbrev == "B cell"] = "Cd3e+ Klrb1+ B cell"
abbrev[grepl("Slamf1-positive",abbrev)] = "Slamf1+ MPC"
abbrev[grepl("Slamf1-negative",abbrev)] = "Slamf1- MPC"
abbrev[grepl("common lymphoid",abbrev)] = "CLP"
abbrev[grepl("megakaryocyte",abbrev)] = "MEP"
abbrev[grepl("hematopoietic",abbrev)] = "HSC"
abbrev[grepl("hematopoietic",abbrev)] = "HSC"
abbrev[grepl("granulocyte monocyte",abbrev)] = "GMP"
# abbrev[grepl("granulocytopoietic",abbrev)] = "G-poiesis"
names(abbrev) = levels(cell_label)
levels(cell_label) = abbrev[levels(cell_label)]

# Create one-hot vector of Cell Type
one_hot_matrix <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
colnames(one_hot_matrix) = unique(cell_label)
rownames(one_hot_matrix) = names(cell_label)

# # Order Cell Type based on its centroid in Featurespace
# Idents(tiss) = cell_label
# ordered_cell_types <- unique(cell_label)[hclust(dist(t(sapply(unique(cell_label), function(x) colMeans(Embeddings(tiss, reduction = "pca")[WhichCells(tiss, idents = x), ,drop = FALSE]) ))))$order]

# Calculate Jaccard index between module activity & cell type
jaccard_index <- apply(cell_block, 2, function(module_col) {
  apply(one_hot_matrix, 2, function(one_hot_col) {
    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
  })
})
# jaccard_index = jaccard_index[rev(ordered_cell_types),]
# rm(ordered_cell_types)
```

### Annotate Each Celltype-specific activated Modules with its corresponding Celltypes
```{r}
module_celltype = apply(jaccard_index,2,function(x){paste(names(which(x > 0.4)),collapse = ",")})
```

## Transferability of Modules
### Load Data & Calculate Module Activity
```{r}
if (!file.exists('droplet_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'droplet_Marrow_seurat_tiss.Robj', method = 'libcurl')
 }
load('droplet_Marrow_seurat_tiss.Robj')
tiss_droplet <- UpdateSeuratObject(tiss_droplet)

# path <- "/data/ruiqi/local_marker/dataset/tabular_muris/"
# tissue_name = "marrow_droplet"
# tiss_droplet <- readRDS(sprintf("%s%s.rds",path,tissue_name))

feature_space_droplet = as.matrix(tiss_droplet@reductions$pca@cell.embeddings[,1:n_dim])
dat_droplet = as.matrix(tiss_droplet[["RNA"]]@data)

knn = 10; major_vote = 5
A = Symmetric_KNN_graph(knn = knn, feature_space = feature_space_droplet)$'adj_matrix'
cell_block_loop = lapply(1:100,function(loop){
  GMM_subsampling(seed = loop, gene_partition, expr_dat = dat_droplet, cell_kNN_graph = A, major_vote)
})
cell_block_prop = Reduce(`+`, cell_block_loop) / length(cell_block_loop)
cell_block_droplet = cell_block_prop
```

### Calculate Jaccard Index
```{r}
cell_label = tiss_droplet$cell_ontology_class
cell_label[!cell_label %in% names(abbrev)] = NA
cell_label = setNames(abbrev[cell_label],names(cell_label))

one_hot_matrix_droplet <- sapply(unique(na.omit(cell_label)), function(x) as.integer(cell_label == x))
one_hot_matrix_droplet[is.na(one_hot_matrix_droplet)] = 0

jaccard_index_droplet <- apply(cell_block_droplet, 2, function(module_col) {
  apply(one_hot_matrix_droplet, 2, function(one_hot_col) {
    1 - philentropy::distance(rbind(module_col,one_hot_col), method = "jaccard", mute.message=TRUE)
  })
})
```
## Save Intermediate data
```{r}
save(gene_partition,dist,gene_hree,cell_block,cell_block_droplet,one_hot_matrix,one_hot_matrix_droplet,jaccard_index,jaccard_index_droplet,file = "./marrow_result/module_result.rdata")
```

## Step4: Module Enrichment Analysis
### GO enrichment Analysis
```{r}
library("clusterProfiler")
library("ReactomePA")
library("org.Mm.eg.db")
library("AnnotationDbi")

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))

# Filter GO terms based on p values < 0.05
ego_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichGO(gene = names(gene_partition)[gene_partition == i],
           OrgDb = 'org.Mm.eg.db', # mouse
           keyType = "SYMBOL",
           ont = "ALL",
           pAdjustMethod = "BH",
           universe = universe_df$symbol,
           pvalueCutoff = 0.05)
})

# Only keep modules which has at lease one GO term having adjusted p-value < 0.05
top_go_clusters <- which(sapply(ego_result, function(x)
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 GO terms
top_enrich_go <- lapply(top_go_clusters, function(i) ego_result[[i]]@result[1:5,])
```

### Reactome enrichment Analysis

```{r}
library("ReactomePA")
library("org.Mm.eg.db")

# Set all genes(after filtering) in the expression matrix as background,
# such that the enrichment result of any module is not attributed to its high expression levels
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
universe_df = universe_df[!is.na(universe_df$entrez),]

# Filter Pathway based on p values < 0.05
epathway_result <- lapply(1:nlevels(gene_partition), function(i){
  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
                organism = "mouse",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05,
                universe = universe_df$entrez)
})

# Only keep modules which has at lease one Pathway having adjusted p-value < 0.05
top_pathway_clusters <- which(sapply(epathway_result, function(x) 
  (x@result$p.adjust[1] < 0.05)))

# Annotate each modules with top5 Pathway terms
top_enrich_pathway <- lapply(top_pathway_clusters, function(i) epathway_result[[i]]@result[1:5,])
```

## Step5: Generate Gene Module Annotation Table
```{r}
df_module = data.frame(Module = levels(gene_partition))
df_module[,"Gene"] = unlist(lapply(levels(gene_partition),function(x){paste(names(gene_partition)[gene_partition == x],collapse = ",")}))
df_module[,"celltype"] = module_celltype[levels(gene_partition)]
df_module[top_go_clusters,"GO_description"] = unlist(lapply(top_enrich_go,function(x){
  des = x[["Description"]]
  paste(des[!is.na(des)],collapse = ",")
}
))
df_module[top_pathway_clusters,"Pathway_description"] = unlist(lapply(top_enrich_pathway,function(x){
  des = x[["Description"]]
  paste(des[!is.na(des)],collapse = ",")
}))
df_module[is.na(df_module)] = ""
write.csv(df_module,file = "./marrow_result/module_description.csv",row.names = FALSE)
```

## Cell Cycle Modules
### Reactome analysis
```{r}
library("ReactomePA")
library("org.Mm.eg.db")
universe_df = data.frame("symbol" = rownames(dat),"entrez" = mapIds(org.Mm.eg.db, keys=rownames(dat), column="ENTREZID", keytype="SYMBOL"))
universe_df = universe_df[!is.na(universe_df$entrez),]
module_cc = c(16,18,17,31,32)
epathway_result_cc <- lapply(module_cc, function(i){
  enrichPathway(gene=universe_df[universe_df$symbol %in% names(gene_partition)[gene_partition == i],"entrez"],
                organism = "mouse",
                pAdjustMethod = "BH",
                pvalueCutoff = 0.05,
                universe = universe_df$entrez)
})
names(epathway_result_cc) = module_cc
# Annotate each modules with top5 Pathway terms
top_enrich_pathway_cc <- do.call(rbind,lapply(1:length(epathway_result_cc), function(i) data.frame(epathway_result_cc[[i]]@result[1:5,],module = module_cc[i])))
top_enrich_pathway_cc <- top_enrich_pathway_cc[complete.cases(top_enrich_pathway_cc), ]
top_enrich_pathway_cc = top_enrich_pathway_cc %>% filter(p.adjust < 0.05)

write.csv(top_enrich_pathway_cc,file = "./marrow_result/reactome_pathway_cc.csv")
```
### Annotate Cell Cycle
```{r}
library(gprofiler2)
mmus_s = gorth(cc.genes.updated.2019$s.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
mmus_g2m = gorth(cc.genes.updated.2019$g2m.genes, source_organism = "hsapiens", target_organism = "mmusculus")$ortholog_name
tiss = CellCycleScoring(tiss, s.features = mmus_s, 
                        g2m.features = mmus_g2m, set.ident = FALSE)
cell_cycle_table = tiss@meta.data[,c("S.Score","G2M.Score","Phase")]
write.csv(cell_cycle_table,file = "./marrow_result/annotated_cell_cycle.csv")
```

<!-- ## Multi-resolution Modules - analysis -->
<!-- ```{r} -->
<!-- df_module = read.csv("./marrow_result/module_description.csv") -->
<!-- df_module$'multi-celltype' = NA -->
<!-- df_module$'note' = NA -->

<!-- ct_prob_per_module = do.call(rbind,lapply(unique(cell_label),function(x){apply(cell_block[cell_label == x,],2,median)})) -->
<!-- rownames(ct_prob_per_module) = unique(cell_label) -->
<!-- ct_num_per_module = sort(apply(ct_prob_per_module,2,function(x)sum(x > 0.5))) -->

<!-- m = 19 -->
<!-- df_module[m,"celltype"] = paste0(names(which(ct_prob_per_module[,m] > 0.5)),collapse = ",") -->

<!-- df_module[54,"multi-celltype"] = c("T/NK/NKT") -->
<!-- df_module[27,"multi-celltype"] = c("Progenitor/NK/NKT") -->
<!-- df_module[33,"multi-celltype"] = c("Progenitor/basophil/granulocytopoietic/half-granulocyte") -->
<!-- df_module[46,"multi-celltype"] = c("monocyte/granulocytopoietic/half-granulocyte") -->
<!-- df_module[48,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte") -->
<!-- df_module[49,"multi-celltype"] = c("monocyte/macrophage/granulocytopoietic/granulocyte") -->
<!-- df_module[55,"multi-celltype"] = c("NK") -->
<!-- df_module[56,"multi-celltype"] = c("NK") -->
<!-- df_module[13,"multi-celltype"] = c("B") -->
<!-- df_module[50,"multi-celltype"] = c("monocyte/granulocytopoietic/granulocyte/basophil") -->
<!-- df_module[28,"multi-celltype"] = c("Progenitor") -->
<!-- df_module[9,"multi-celltype"] = c("immature/naive B") -->
<!-- df_module[11,"multi-celltype"] = c("late pro/pre B") -->
<!-- df_module[7,"multi-celltype"] = c("late pro/pre B") -->
<!-- df_module[44,"multi-celltype"] = c("granulocytopoietic/half-granulocyte") -->
<!-- df_module[47,"multi-celltype"] = c("granulocytopoietic/granulocyte") -->
<!-- df_module[52,"multi-celltype"] = c("T") -->
<!-- df_module[53,"multi-celltype"] = c("macrophage/NK") -->
<!-- df_module[1,"multi-celltype"] = c("monocyte/macrophage/granulocyte monocyte progenitor cell") -->
<!-- df_module[20,"multi-celltype"] = c("granulocytopoietic/Progenitor") -->
<!-- df_module[c(42,35),"multi-celltype"] = c("below-half-granulocyte/monocyte") -->
<!-- df_module[c(36,38),"multi-celltype"] = c("granulocyte monocyte progenitor cell/monocyte") -->
<!-- df_module[2,"multi-celltype"] = c("granulocyte monocyte progenitor cell/macrophage") -->
<!-- df_module[4,"celltype"] = c("granulocyte monocyte progenitor cell") -->
<!-- df_module[c(5,6,43),"note"] = "subtype" -->

<!-- df_module[c(16,18,17,31,32),'note'] = "cell_cycle" -->
<!-- ``` -->