### 2024-03-26 Marker_identfication_benchmark.Rmd
24/03/26

---
title: "Marker_identification_Benchmark"
author: "Ruiqi Li"
output: pdf_document
date: '2023-10-16'
---

# Load Packages
```{r}
if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")
if (!requireNamespace("BiocManager", quietly = TRUE)) BiocManager::install("Seurat")
if (!requireNamespace("readxl", quietly = TRUE)) install.packages("readxl")
if (!requireNamespace("dplyr", quietly = TRUE)) install.packages("dplyr")

library("Seurat")
library("dplyr")
library(RColorBrewer)
```

# Load Data & Process
## Load Data
```{r}
# tiss <- readRDS("/data/ruiqi/local_marker/dataset/tabular_muris/marrow_facs.rds")
# DefaultAssay(tiss) <- "RNA"
```

```{r}
tissue_fullname = "Bone marrow"

if(tissue_fullname == "Bone marrow"){
  tissue_name = "marrow_facs"
  if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Marrow_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Pancreas"){
  tissue_name = "pancreas_facs"
  if (!file.exists('facs_Pancreas_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092386",
                  destfile = 'facs_Pancreas_seurat_tiss.Robj', method = 'libcurl')
  }
  load('facs_Pancreas_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Lung"){
  tissue_name = "lung_facs"
  if (!file.exists('facs_Lung_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://figshare.com/ndownloader/files/13092194",
                  destfile = 'facs_Lung_seurat_tiss.Robj', method = 'libcurl')
   }
  load('facs_Lung_seurat_tiss.Robj')
  tiss <- UpdateSeuratObject(tiss)
}
if(tissue_fullname == "Thymic"){
  options(timeout=6000)
  if (!file.exists('Human_Thymic_Epithelial_seurat_tiss.Robj')) {
     # If load url doesn't work, please download it and load from your working directory
    options(timeout=6000)
    download.file("https://datasets.cellxgene.cziscience.com/046d2e35-2ee6-4159-8db7-7b1039fe0005.rds",
                  destfile = 'Human_Thymic_Epithelial_seurat_tiss.Robj', method = 'libcurl')
  }
  load('Human_Thymic_Epithelial_seurat_tiss.Robj')
}
```

## Process
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = pmin(dim(tiss@reductions$pca@cell.embeddings)[2],30)
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)

Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
selected_genes = names(selected_genes)[selected_genes]
dat = dat[selected_genes,,drop = FALSE]
raw_count = as.matrix(tiss@assays$RNA@counts)[selected_genes,,drop = FALSE]
total_count = tiss@meta.data$nCount_RNA
```

# Define Ground truth Marker
## Criterion1: Fold-change
For each gene, sorting cell types by mean expression for each gene and computing fold changes between consecutive types, take the maximum value among the N-1 fold change values, given N cell types.
```{r}
folder_path = "benchmark_result"
# Check if the folder exists
if (!dir.exists(folder_path)) {
  # If it doesn't exist, create it
  dir.create(folder_path)
}
```

```{r}
avg_exp = AverageExpression(subset(tiss,features = selected_genes), assays = "RNA", slot = "counts", group.by = "cell_ontology_class") %>% as.data.frame()

avg_exp_ordered <- avg_exp %>%
  rowwise() %>%
  mutate(cell_order_val = list(sort(c_across(everything()), decreasing = FALSE))) %>%
  ungroup()

avg_exp_ordered = avg_exp_ordered %>%
  rowwise() %>%
  mutate(log_fold_change = list(c(NA, diff(log2(unlist(cell_order_val)+1))))) %>%
  ungroup()

max_logfc = unlist( lapply(avg_exp_ordered$log_fold_change,function(vec){max(vec,na.rm = TRUE)}) )
names(max_logfc) = rownames(avg_exp)

max_logfc = sort(max_logfc,decreasing = TRUE)
write.table(max_logfc,file = file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt")))
```
## Criterion2: CellMarkerDB
Downloaded marker from 
http://117.50.127.228/CellMarker/CellMarker_download_files/file/Cell_marker_Seq.xlsx
```{r}
# Load CellMarkerDB
cell_marker_db = readxl::read_xlsx("/data/ruiqi/local_marker/dataset/Cell_Marker/Cell_marker_All.xlsx") %>% filter(species == "Mouse") %>% filter(tissue_class == "Bone marrow") %>% filter(!is.na(cellontology_id)) %>% filter(!is.na(PMID)) %>% mutate(gene_name1 = marker) %>% mutate(gene_name2 = Symbol) %>% select(c("gene_name1","gene_name2")) 
# Lung, Pancreas

# Match gene name
celldb_marker = union(cell_marker_db$gene_name1,cell_marker_db$gene_name2)
celldb_marker = unlist( lapply(celldb_marker,function(i){
  id = grep(paste0("^", i, "$"), selected_genes, ignore.case = TRUE)
  if(length(id)==0){return(NA)}
  else{return(selected_genes[id])}
}) )
celldb_marker = celldb_marker[!is.na(celldb_marker)] %>% unique()
write.table(celldb_marker,file = file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))
```

# Run Each Method and save results
## LocalMarkerDetector
Genes were ranked by the increasing order of Cumulative Diffusion-KL score
```{r}
method = "lmd"
source("LMD_function.R", echo = F)
res = LMD(dat,feature_space,max_time = 2^20,knn = 5)
marker = show_result_lmd(res, n = length(selected_genes))
marker[order(marker$score),'rank'] = 1:nrow(marker)
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Seurat v4 Wilcoxon unfiltered
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)
method = "wilcox_no_filter"
# marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0, min.pct = 0)
# gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
marker <- presto:::wilcoxauc.Seurat(X = subset(tiss,features = selected_genes),
                                    assay = "data", 
                                    seurat_assay = "RNA")
colnames(marker)[1] = "gene"
gene_rank = marker %>% arrange(.,padj,desc(logFC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```
## Seurat v4 default
Genes were ranked by their lowest adjusted P-value in any cluster (increasing order), if two genes had the same, prioritize genes with larger FC.
```{r}
tiss <- FindNeighbors(tiss, dims = 1:n_dim)
tiss <- FindClusters(tiss, resolution = 1.5, algorithm = 3)

method = "wilcox_default"
marker <- FindAllMarkers(tiss, features = selected_genes, only.pos = FALSE, test.use = "wilcox", logfc.threshold = 0.25,min.pct = 0.1)

gene_rank = marker %>% arrange(.,p_val_adj,desc(avg_log2FC)) %>% distinct(.,gene, .keep_all = TRUE) %>% select(gene)
gene_rank$'rank' = 1:nrow(gene_rank)
marker <- merge(marker,gene_rank,by = "gene")
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```
## singleCellHaystack
https://alexisvdb.github.io/singleCellHaystack/articles/examples/a02_example_scRNAseq.html
Genes were ranked based on the increasing order of log(adjusted P-value)
```{r}
method = "haystack"
library(singleCellHaystack)
set.seed(123)
res <- haystack(x = feature_space, expression = dat)
marker <- show_result_haystack(res.haystack = res)
marker[order(marker$log.p.adj),"rank"] = 1:nrow(marker)

write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```
## SEMITONES
https://github.com/ohlerlab/SEMITONES/tree/master
https://github.com/ohlerlab/SEMITONES_paper/blob/main/notebooks/10_SEMITONES_benchmark_corrected.ipynb

Genes were ranked by their lowest adjusted P-value in any reference cell (increasing order), if two genes had the same p.adj, prioritize genes with larger absolute enrichment score in any reference cell.
```{python}
import SEMITONES
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
from SEMITONES.cell_selection import from_knn_dist
from SEMITONES.cell_selection import from_gui
from SEMITONES.cell_selection import from_2D_embedding
from SEMITONES.cell_selection import get_cells_from_gui
from sklearn.metrics.pairwise import pairwise_kernels
from SEMITONES.enrichment_scoring import calculate_escores
from SEMITONES.enrichment_scoring import permute
from SEMITONES.enrichment_scoring import sig_interval
from SEMITONES.support_funcs import pairwise_similarities
from SEMITONES.support_funcs import sig_dictionary
from SEMITONES.support_funcs import sig_bool
from SEMITONES.enrichment_scoring import pvals_per_cell

# Load data
data = pd.DataFrame(r.dat.T)  # counts
PC20 = pd.DataFrame(r.feature_space)
# Reference Cell
g = 8.6e-4  
S = pairwise_kernels(PC20, metric="rbf", gamma=g) 
median = np.median(S, axis=0)
start = int(np.argmin(median))
num_refcell = int(data.shape[1]*0.005)
dd_rcells = from_knn_dist(X=PC20, 
                          n_ret=num_refcell,
                          start=start, 
                          metric="rbf",
                          metric_params={"gamma": g})
S = pairwise_similarities(PC20,
                          query=dd_rcells,
                          metric="rbf",
                          metric_params={"gamma": g})
dd_rcells = data.index[dd_rcells]
escores = calculate_escores(data.values, query=dd_rcells, S=S)

# semi_ranks = escores.abs().rank(ascending=False, method="dense").min(1)

P = permute(data.values)
pscores = calculate_escores(P, query=dd_rcells, S=S)
pvals = pvals_per_cell(escores,pscores,ret = "q")
# semi_ranks = pvals.min(1).rank(ascending=True, method="dense")

result = pd.DataFrame({'escore': escores.abs().max(1), 'padj': pvals.min(1)})
```
```{r}
method = "semi"
marker = reticulate::py$result
rownames(marker) = rownames(dat)
marker = marker %>% arrange(.,padj,desc(escore))
marker$'rank' = 1:nrow(marker)

write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Hotspot
https://hotspot.readthedocs.io/en/latest/CD4_Tutorial.html
Genes were ranked by the increasing order FDR value (if two genes had the same, prioritize genes with larger Z-score).
```{r}
cell_id = colnames(raw_count)
gene_id = rownames(raw_count)
```
```{python}
import hotspot
import scanpy as sc
import pandas as pd
from scipy.sparse import csc_matrix
adata = sc.AnnData(X = r.raw_count.T,obs = pd.DataFrame(index = r.cell_id), var = pd.DataFrame(index = r.gene_id))
adata.obs['total_counts'] = r.total_count
adata.layers["counts"] = adata.X.copy()
adata.obsm['X_pca'] = r.feature_space
adata.layers["counts_csc"] = csc_matrix(adata.layers["counts"])

hs = hotspot.Hotspot(
    adata,
    layer_key="counts_csc",
    model='danb',
    latent_obsm_key="X_pca",
    umi_counts_obs_key="total_counts"
)
hs.create_knn_graph(weighted_graph=False, n_neighbors=30)

# Determining informative genes
hs_results = hs.compute_autocorrelations()
```
```{r}
method = "hotspot"
marker = reticulate::py$hs_results
marker[order(marker$FDR),"rank"] = 1:nrow(marker)
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Marcopolo
Long Runtime Option
```{r}
# transfer the seurat obj to h5ad & process in jupyter notebook
library(reticulate)
library(anndata)
sc <- import("scanpy", convert = FALSE)
scvi <- import("scvi", convert = FALSE)
# scvi$settings$progress_bar_style = 'tqdm'
tiss_adata <- sc$AnnData(
  X = t(raw_count)
)
anndata::write_h5ad(tiss_adata, "./dataset/tm_marrow.h5ad")
```
The rank of genes is given by `Marcopolo_rank`
```{python}
import pickle
import numpy as np
import pandas as pd
import torch
import anndata as ad
import scanpy as sc
import matplotlib.pyplot as plt
import MarcoPolo

adata = sc.AnnData(X = r.raw_count.T)

if "size_factor" not in adata.obs.columns:
    norm_factor = sc.pp.normalize_total(adata, exclude_highly_expressed=True, max_fraction= 0.2, inplace=False)["norm_factor"]
    adata.obs["size_factor"] = norm_factor/norm_factor.mean()
    print("size factor was calculated")

regression_result = MarcoPolo.run_regression(adata=adata, size_factor_key="size_factor",
                         num_threads=1, device="cpu")

result = MarcoPolo.find_markers(adata=adata, regression_result=regression_result)
result = result.set_axis(list(adata.var.index) ,axis = 0)
```
```{r}
method = "marcopolo"
marker = reticulate::py$result
marker$'rank' = marker$MarcoPolo_rank + 1
write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = TRUE)
```

## Highly Variable Genes
```{r}
method = "hvg"

DefaultAssay(tiss) <- "RNA"
# tiss <- NormalizeData(tiss, normalization.method = "LogNormalize", scale.factor = 1e6)
marker <- subset(tiss,features = selected_genes) %>% ScaleData(features = selected_genes) %>% 
  FindVariableFeatures(nfeatures = 5000) %>% VariableFeatures() %>% as.data.frame() %>% setNames(c("gene")) %>% mutate(rank = 1:n())

write.table(marker,file = file.path(folder_path, paste0(method,"_",tissue_name,".csv")),row.names = FALSE)
```

# Integrate rank from all methods into one dataframe
```{r}
method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
df_benchmark = lapply(method_ls, function(method){
  marker = read.table(file.path(folder_path, paste0(method,"_",tissue_name,".csv")),header = TRUE)
  if(!"gene" %in% colnames(marker)){
    marker$'gene' = rownames(marker)
  }
  marker %>% select(gene,rank) %>% distinct()
})
names(df_benchmark) = method_ls
df_benchmark = bind_rows(df_benchmark, .id = "method")
df_benchmark <- tidyr::pivot_wider(df_benchmark, names_from = method, values_from = rank)
df_benchmark[is.na(df_benchmark)] = nrow(df_benchmark)

# rownames(df_benchmark) = df_benchmark$gene
# df_benchmark[order(Gene_detected_count[rownames(df_benchmark)]),"gene_count"] = 1:nrow(df_benchmark)

write.table(df_benchmark,file = file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")),row.names = FALSE)
```

# AUROC
```{r}
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))

method_ls = c("lmd","haystack","hotspot","marcopolo","semi","wilcox_default","wilcox_no_filter","hvg")
```
```{r}
max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]
gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

# AUC table
df = do.call(rbind,lapply(1:length(gt_list),function(i){
  true_marker = gt_list[[i]]
  df_benchmark$"gt" = 0
  df_benchmark[true_marker,"gt"] = 1

  library(pROC)
  roc_df = do.call(rbind,lapply(setdiff(colnames(df_benchmark),"gt"), function(method){
    roc = roc(df_benchmark$gt, df_benchmark[,method], direction = ">")
    data.frame(FPR = 1 - roc$specificities, TPR = roc$sensitivities, Method = method, AUROC = as.numeric(auc(roc)))
  }) )

  df = roc_df %>% select(c("Method","AUROC")) %>% distinct()
  df$'gt_set' = names(gt_list)[i]
  df
}))
df$Method = as.factor(df$Method)
df$gt_set = factor(df$gt_set,levels = names(gt_list))
write.table(df,file = file.path(folder_path, paste0(tissue_name,"_auroc.csv")),row.names = FALSE)
```

```{r}
# Concordance table
do.call(cbind,lapply(gt_list[c(2,21)],function(true_marker){
  df_benchmark$"gt" = 0
  df_benchmark[true_marker,"gt"] = 1
  apply(df_benchmark,2,function(x){
    sum(df_benchmark[order(x),"gt"][1:100])
  })
}))

# AUC curve
p = ggplot(roc_df, aes(x = FPR, y = TPR, color = label)) +
  geom_line(size = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray") + labs(title = sprintf("CellMarkerDB (%d genes)",length(true_marker)),
     x = "False Positive Rate",
     y = "True Positive Rate") +  scale_colour_manual(values= coldef)
```

# Density Index
```{r}
folder_path = "benchmark_result"
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    # top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
    top_n = c(seq(50,90,10),seq(100,1000,100))
  }
  DI = unlist(lapply(top_n, function(top){
    features = top_gs[1:top]
    # remove drop-out cells
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
    # # Silhouette score
    # d = dist(Embeddings(subtiss, reduction = "pca"))
    # c = as.integer(factor(subtiss$cell_ontology_class[labels(d)]))
    # s1 = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
    #   group_by(cluster) %>%
    #   summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # Density-index
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    s2 = length_scale/mean(nn.dist)

    s2

  }))
  # rownames(result) = c("Mean_SI","DI")

  # Non-dropout Rate
  cell_count = lapply(top_n,function(n){
    colSums(dat[top_gs[1:n],]>0)
  })
  non_dropout_rate = unlist(lapply(cell_count,function(x){
    sum(x != 0)/length(x)
  }))

  result = data.frame(TopGene = top_n,Method = method,DensityIndex = DI, non_dropout_rate = non_dropout_rate)
  result
}) )

write.table(score_measure,file = file.path(folder_path, paste0(tissue_name,"_density_index_no_dropout.csv")),row.names = FALSE)

# error bar & null distribution
top_n = c(seq(50,90,10),seq(100,900,100),seq(1000,3000,500),seq(4000,9000,1000),seq(10000,max(df_benchmark),2000))
null_score_measure = do.call(rbind,lapply(top_n,function(top){
  set.seed(top)
  seed_ls = sample(1:1e5,size = 100)
  random_top_gs = lapply(seed_ls,function(sed){
    set.seed(sed)
    sample(rownames(df_benchmark),top,replace = FALSE)
  })
  result = unlist(lapply(random_top_gs, function(g_ls){
    features = g_ls
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)
    # Density-index
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2))
    s2 = length_scale/mean(nn.dist)
    s2
  }))
  result = data.frame(TopGene = top,DensityIndex = result)
}) )
write.table(null_score_measure,file = file.path(folder_path, paste0(tissue_name,"_DI_null.csv")),row.names = FALSE)

# smallest set of genes covering 95% cells & corresponding DI
min_coverage = 1
coverage_score_measure = do.call(rbind,lapply(colnames(df_benchmark),function(method){
  cat(method,"\n")
  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
  min_covered = apply(dat[top_gs,] > 0, 2, function(x) which(x)[min_coverage])
  min_row_for_coverage <- as.integer(names(which(cumsum(table(min_covered)) >= ncol(dat) * 0.95)[1]))
  features = top_gs[1:min_row_for_coverage]

  # remove drop-out cells
  subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
  npcs = pmin(20,length(features))
  subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = npcs, verbose = FALSE)
  subtiss <- RunUMAP(subtiss,dims = 1:dim(subtiss@reductions$pca)[2])
  # Density-index
  if(length(features) < npcs){
    nn.dist = FNN::get.knn(t(subtiss[["RNA"]]@scale.data), k = 10, algorithm = "kd_tree")$nn.dist
    length_scale <- norm(t(subtiss[["RNA"]]@scale.data),type = "F")/sqrt(ncol(subtiss)-1)
    s2 = length_scale/mean(nn.dist)
  }else{
    nn.dist = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")$nn.dist
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    s2 = length_scale/mean(nn.dist)
  }

  data.frame(TopGene = min_row_for_coverage,Method = method,DensityIndex = s2, non_dropout_rate = sum(colSums(dat[features,]>0)!=0)/ncol(dat))

  DimPlot(subtiss,group.by = "cell_ontology_class") + NoLegend() + ggtitle(sprintf("%s, %d genes, DI: %.4f",method,min_row_for_coverage,s2))
}))

```

# NNret + Silhouette + DI
```{r}
folder_path = "benchmark_result"
df_benchmark = read.table(file.path(folder_path, paste0(tissue_name,"_benchmark_rank_table.csv")))
score_measure =do.call(rbind,lapply(c(colnames(df_benchmark),"All genes"),function(method){
  if(method == "All genes"){
    top_gs = rownames(df_benchmark)
    top_n = length(top_gs)
  }else{
    top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
    top_n = c(seq(50,90,10),seq(100,1000,100))
  }
  scores = sapply(top_n, function(top){
    features = top_gs[1:top]
    # remove drop-out cells
    subtiss = subset(tiss, cells = which(colSums(subset(tiss,features = features))>0))
    subtiss <- subtiss %>% ScaleData(features = features) %>% RunPCA(features = features, npcs = 20)

    # Silhouette score
    d = dist(Embeddings(subtiss, reduction = "pca"))
    c = as.integer(factor(subtiss$celltype[labels(d)]))
    si = mean(as.data.frame(cluster::silhouette(c, d)[,]) %>%
      group_by(cluster) %>%
      summarize(mean_silhouette = mean(sil_width)) %>%.$mean_silhouette)

    # NNret
    nn.result = FNN::get.knn(Embeddings(subtiss, reduction = "pca"), k = 10, algorithm = "kd_tree")
    nn.dist = nn.result$nn.dist
    nn.index = nn.result$nn.index
    nnret = mean(
      unlist(lapply(1:nrow(nn.index),function(i){
      mean(c[nn.index[i,]] == c[i])
    })) )

    # Density-index
    sdVec <- stats::na.omit(subtiss@reductions$pca@stdev)
    length_scale <- sqrt(sum(sdVec^2)) # length-scale * sqrt(2*(N-1)/N) = d_rms
    di = length_scale/mean(nn.dist)

    c(si,di,nnret)
  }) 
  scores = t(scores)
  colnames(scores) = c("Sihouette_score","DensityIndex","NNret")
  rownames(scores) = top_n

  # Non-dropout Rate
  cell_count = lapply(top_n,function(n){
    colSums(tiss[["RNA"]]@data[top_gs[1:n],]>0)
  })
  non_dropout_rate = unlist(lapply(cell_count,function(x){
    sum(x != 0)/length(x)
  }))

  result = data.frame(TopGene = top_n,Method = method, scores, non_dropout_rate = non_dropout_rate)
  result
}) )
```

# Recovery of celltypes
```{r}
# jaccard index of cell-type-specificity of each genes
cell_label = tiss$celltype
one_hot_matrix_ct <- sapply(unique(cell_label), function(x) as.integer(cell_label == x))
binary_dat = dat > apply(dat,2,median)
intersect_mtx = binary_dat %*% one_hot_matrix_ct
union_mtx = sapply(colnames(one_hot_matrix_ct),function(ct){
  rowSums((binary_dat > 0)) + colSums(one_hot_matrix_ct)[ct]
}) - intersect_mtx
jaccard_sim = intersect_mtx / union_mtx
threshold = 0.8 / ((ncol(tiss)/table(cell_label))*0.2 + 0.8)[colnames(jaccard_sim)]
recovery_mtx = t(apply(jaccard_sim,1,function(x) x > threshold))

# Prop of recovered cell types
recovery_rate =data.frame(sapply(c(colnames(df_benchmark)),function(method){
  top_gs = rownames(df_benchmark)[order(df_benchmark[,method])]
  top_n = 1:300
  scores = sapply(top_n, function(top){
    features = top_gs[1:top]
    mean(colSums(recovery_mtx[features,,drop = FALSE]) > 0)
  }) 
}) )
recovery_rate$'TopGene' = 1:nrow(recovery_rate)
```

# Sensitivity Test (Try different K)
```{r}
# RunLMD under differen K
source("./LMD_function.R", echo = F)
score_df = do.call(cbind,lapply(c(3:10,20,30,40,50),function(kNN){
  try({
    res = LMD(dat,feature_space,knn = kNN)
    return(res$cumulative_score)
  },silent = TRUE)
}))
colnames(score_df) = paste0("kNN",c(3:10,20,30,40,50))
write.table(score_df,file = file.path("marrow_result", paste0(tissue_name,"_different_knn.csv")),row.names = TRUE)
```
```{r}
# Obtain AUROC
score_ranked <- data.frame(score_df) %>% mutate(across(everything(), rank))

max_logfc = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c1.txt"))) %>% rownames()
celldb_marker = read.table(file.path(folder_path, paste0(tissue_name,"_ground_truth_c2.txt")))[,1]

gt_list = c(lapply(seq(50,1000,50),function(x) max_logfc[1:x]),list(celldb_marker))
names(gt_list) = c(paste0("Top",seq(50,1000,50)),"CellMarkerDB")

auc_df = apply(score_ranked,2,function(x){
  unlist(lapply(gt_list,function(genes){
    roc = pROC::roc(as.integer(rownames(score_ranked) %in% genes), x, direction = ">")
    pROC::auc(roc)
  }))
})

auc_df = data.frame(auc_df,gt = names(gt_list))
write.table(auc_df,file = file.path("marrow_result", paste0(tissue_name,"_sensitivity_auroc.csv")),row.names = FALSE)
```