### 2024-05-23 Demo_marrow_facs.Rmd
240523 pipeline-development

---
title: "Tutorial for Runing LMD on Mouse Bone Marrow Dataset"
author: "Ruiqi Li"
date: '2023-10-12'
output: pdf_document
bibliography: references.bib
---

# Install & Load Packages
```{r}
source("LMD_function.R")
# Define package lists
bioc_packages <- c("clusterProfiler", "AnnotationDbi", "ReactomePA", "org.Mm.eg.db","gprofiler2")
github_packages <- c("satijalab/seurat-wrappers");
names(github_packages) = strsplit(github_packages, "/")[[1]][2]
names(github_packages)[grepl("seurat-wrappers",names(github_packages))] = "SeuratWrappers"
# Install packages
sapply(bioc_packages, function(pkg) if (!requireNamespace(pkg, quietly = TRUE)) BiocManager::install(pkg))
sapply(1:length(github_packages), function(i) if (!requireNamespace(names(github_packages)[i], quietly = TRUE)) remotes::install_github(github_packages[i]))
lapply(c(bioc_packages,names(github_packages)), require, character.only = TRUE)
```

# Create a folder to save results
```{r}
folder_path = "marrow_result"
if (!dir.exists(path)) {
  # If it doesn't exist, create it
  dir.create(path)
} else {
}
```

# Preparing input data
```{r}
# tiss <- readRDS("/banach1/ruiqi/local_marker/LMD_data/tabular_muris/marrow_facs.rds")
```

```{r}
if (!file.exists('facs_Marrow_seurat_tiss.Robj')) {
   # If load url doesn't work, please download it and load from your working directory
  options(timeout=6000)
  download.file("https://figshare.com/ndownloader/files/13092380",
                destfile = 'facs_Marrow_seurat_tiss.Robj', method = 'libcurl')
 }
load('facs_Marrow_seurat_tiss.Robj')
tiss <- UpdateSeuratObject(tiss)
```
We obtain the following objects from this Seurat Obj:
* `feature_space`: The output of PCA. This matrix contains the first 20 principal components (PCs).
* `visual_space`: a data.frame with t-SNE coordinates (2D) based on the first 20 PCs.
* `dat`: a matrix object with the log-normalized expression of genes (rows) in each cell (columns)
* `cell_label`: Any metadata you wish to visualize and examine related to cells (here we take the cell type annotation)
```{r}
DefaultAssay(tiss) <- "RNA"
n_dim = dim(tiss@reductions$pca@cell.embeddings)[2]
feature_space = as.matrix(tiss@reductions$pca@cell.embeddings[,1:n_dim])
visual_space = data.frame(tiss@reductions$tsne@cell.embeddings)
dat = as.matrix(tiss[[DefaultAssay(tiss)]]@data)
cell_label = tiss$cell_ontology_class
```

We process the `dat` by only keep the genes detected in more than 10 cells and less than 50% of cells (genes is "detected" in a cell only if its expression level is greater than the median expression level of that cell)

```{r}
Gene_detected_count <- apply(dat > apply(dat,2,median),1,sum)
selected_genes = (Gene_detected_count >= 10) & (Gene_detected_count <= ncol(dat) * 0.5)
dat = dat[selected_genes,,drop = FALSE]
```

```{r}
# # save to hdf5 file
# df = sub.srat@meta.data[,c("design_coarse","celltype1205")]
# # BiocManager::install("rhdf5")
# library(rhdf5)
# h5write(dat, 'Bone_marrow_facs_mtx.h5', 'dataset')
# fwrite(df,file = "HL234_meta.txt",row.names = TRUE)
```

# Running LMD Step by Step

This provides a step-by-step tutorial for a better understanding of the algorithm. If you want to generate results directly, please skip to 'Running LMD in One Step'

```{r}
source("LMD_function.R", echo = F)
```

## Step1: Build KNN Graph

```{r, fig.width=20}
# Construct knn graph
knn_result = Symmetric_KNN_graph(knn = 5, feature_space = feature_space)

A = knn_result$adj_matrix # Adjacency Matrix
W = knn_result$graph # Symmetrized Graph ((A + AT) / 2)

# Plot knn graph
plot_knn_graph(affinity_m = W, label = cell_label, layout = visual_space)
```

## Step2: Diffuse Genes on the cell graph

```{r, fig.width=20}
# Generate $P^{t}, t = 2,4,8,...,max_time$
P_ls = Obtain_Pls(W = W, max_time = 2^10)
# Generate initial gene distribution
rho = Rowwise_normalize(dat[,colnames(W)])
# Check One Gene Diffusion on Cell graph by time
gene = "Fcnb"
print(FeaturePlot_diffusion(coord = visual_space,init_state = rho[gene,],P_ls = P_ls,check_time = c(2,16,128),gene_name = gene))
```

## Step3: Obtain Diffusion KL score & Cumulative score

```{r}
# Obtain diffusion KL score
score_result = fast_get_lmds(W = W, init_state = rho, P_ls = P_ls)

# Check Diffusion KL pattern for several genes
genes = c("Tlr11","Fcnb","Cd79a","4933425H06Rik","Uchl4","Dock11")
genes_label = rep(c("good","bad"),each = 3)
Visualize_score_pattern(score_result$'score_profile', genes = genes, label_class = genes_label, facet_class = NULL)

# Show result
show_result_lmd(score_result,n = 10)
```

# Running LMD in one step
```{r}
tic()
lmd_result = LMD(dat, feature_space)
toc()
# lmd_result = LMD(dat, feature_space, highres = TRUE, max_time = 2^7)
```

# Identify Gene Modules

## Step1: Deriving Gene Modules
```{r}
# Select top1000 genes as local genes
local_gene = show_result_lmd(lmd_result, n = 1000)
local_gene = rownames(local_gene)

# Denoising
# We use ALRA @linderman2022zero to Denoise our data
tiss <- RunALRA(tiss, assay = "RNA")
DefaultAssay(tiss) <- "RNA"
dat_alra = as.matrix(tiss[["alra"]]@data)[local_gene,]

# Calculate Gene-gene co-localization distance
distance_method = "jaccard"
dist = Calculate_distance(dat_alra, method = distance_method)

# Obtain Gene modules
gene_partition = Obtain_gene_partition(dist, clustering_method = "average", min_gene = 5)
```

### Visualize Heatmap
```{r}
p = Visualize_gene_heatmap(dist)
```

## Step2: Computing Per-Cell Module Activity Scores
```{r}
tiss = AddModuleActivityScore(tiss, gene_partition = gene_partition, do_local_smooth = TRUE)
```