### 2025-04-21 README.Rmd
250421

---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures",
  out.width = "100%"
)
```

# Localized Marker Detector

<!-- badges: start -->
<!-- badges: end -->

Localized Marker Detector (LMD) is a computational framework designed for the identification of gene expression markers localized to specific cell populations within single-cell RNA sequencing data. The major workflow of LMD comprises the following three main steps:

* Step1. Constructing a cell-cell affinity graph
* Step2. Diffusing the gene expression value across the cell graph
* Step3. Assigning a score to each gene based on the dynamics of its diffusion process
* Optional Downstream tasks
  * Identifying gene modules and characterizing functional cell groups
  * Cross-sample comparison

```{r pressure, echo=FALSE, out.width="100%", out.height="auto"}
knitr::include_graphics("./man/figures/LMD_workflow.png")
```

## Installation

LMD can be installed in R as follows:

``` r
install.packages("devtools")
devtools::install_github("ruiqi0130/LocalizedMarkerDetector")

library("LocalizedMarkerDetector")
```

### Docker installation

Alternatively, we provide an option of installing LocalizedMarkerDetector with the accompanied lmd docker environment. The pre-built docker image can be downloaded using

``` bash
docker pull ruiqi0130/lmd:lastest
```

and to run the docker container:

``` bash
docker run -it --rm -e USER=john -e USERID=$(id -u) --name container_name -p 8029:8787  -v /data/:/data/   docker.io/jyc34/bipca-experiment:lastest
```

Here, change `/data:/data to `<your_local_data_directory>:/data`. A rstudio will be on port 8029. If you would like to link a local bipca installation (for package development, for instance), you can write -v <your_local_bipca_directory:/bipca.

## Example tutorial
Please check [LMD tutorial](https://ruiqi0130.github.io/LocalizedMarkerDetector/articles/).

## References
References of LMD functions can be found [here](https://ruiqi0130.github.io/LocalizedMarkerDetector/reference/index.html).

<!-- Data used in this tutorial can be downloaded from [Tabula Muris](https://figshare.com/ndownloader/files/13092380). -->